THis file contains the data processing and analysis required in peer assignment 1  
========================================================================================

The data we are processing is of that of a activity monitoring device of an anonymous individual over a period  
two months

### loading and processing the data

THe data is in the csv format we use to following code to read the data

```{r loading the data}
activity_data<-read.csv("activity.csv")
summary(activity_data)
```

### Mean of total number of steps taken per day
for this we shall ignore the NAs in the dataset and consider only the complete cases
```{r data processing 1}
complete_cases<-activity_data[complete.cases(activity_data),]
```

we shall use the data table package to find out the mean and median of total steps per day and also draw the histogram of total steps taken per day

```{r find the mean per day,fig.height=4}
install.packages("data.table")
library(data.table)
dt<-data.table(complete_cases)
sumsteps_perday<-dt[,list(sum=sum(steps)),by=date]
hist(sumsteps_perday$sum,col="red",xlab="total steps per day")
meansteps_perday<-dt[,list(mean=mean(steps),median=median(steps)),by=date]
summary(meansteps_perday)
```

### plot of time intervel and average number of steps

```{r}
library(lattice)
dt<-data.table(complete_cases)
averagesetps_perintervel<-dt[,mean(steps),by=interval]
attach(averagesetps_perintervel)
xyplot(V1~interval,type="l",ylab="average number of steps")
```````

## Inputing missing values
### Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
The number of rows with missing values is:

```{r report of NAs}
incomplete_cases<-activity_data[!complete.cases(activity_data),]
dim(incomplete_cases)[1]
```
### Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
The NAs are replaced by mean steps of that intervel
```{r replaces NAs,cache=TRUE}
processed_data<-activity_data
for(i in 1:nrow(processed_data)){
    row<-processed_data[i,]
    if(is.na(row$steps)){
      processed_data[i,"steps"]<-averagesetps_perintervel[averagesetps_perintervel$interval==row$interval,V1]
    }
    i<-i+1
  }
```
### now it can be checked that every row is a complete case  
```{r check for NA elimination}
incomplete_cases<-activity_data[!complete.cases(activity_data),]
dim(incomplete_cases)[1]
```

### Make a histogram of the total number of steps taken each day  
```{r histogram of steps per day}
library(data.table)
dt<-data.table(processed_data)
new_sumsteps_perday<-dt[,sum(steps),by=date]
hist(new_sumsteps_perday$V1,col="red",xlab="total steps per day")
```

###Calculate and report the mean and median total number of steps taken per day

```{r recalculate mean and medians}
new_meansteps_perday<-dt[,list(mean=mean(steps),median=median(steps)),by=date]
summary(new_meansteps_perday)
```
### the impace seems to be very minimal. THe median values are always coming as Zeros irrespective of processing
```{r create weekday weekend factor}
processed_data$weekday<-"weekday"
for(i in 1:nrow(processed_data)){
    row<-processed_data[i,]
    if(weekdays(as.Date(row$date)) %in% c("Saturday","Sunday")){
      processed_data[i,"weekday"]<-"weekend"
    }
    i<-i+1
  }
```